---
title: "LEGEND -- Hypertension Phenotypes"
bibliography: CohortDefinition.bib
link-citations: yes
output:
  bookdown::html_document2: default
csl: nature-publishing-group-vancouver.csl 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

# Definition process {-}

Outcomes for LEGEND were initially identified from endpoints from clinical trials reporting in clinical guidelines and systematic reviews, and augmented by adverse events listed on US structured product labels of hypertension drugs (https://dailymed.nlm.nih.gov/dailymed/).
Each outcome was then assessed to determine if it was evaluable in observational data by developing and evaluating an operational phenotype definition.
All phenotypes were designed, implemented and evaluated using the same approach:  Initial design was guided by literature review in PubMed of prior observational studies of administrative claims or electronic health records where the phenotype was used as a study outcome, and if available, validated through source record verification or other means.
Literature review was complemented by review of eMERGE PheKB phenotype entries (https://phekb.org/phenotypes).
Clinical definitions of cardiovascular outcomes were determined from clinical guidelines and systematic review of clinical trials of hypertension treatments (@Reboussin2018 @Whelton2018 @Williams2018).
Where possible, conceptsets originated with published codelists (e.g. ICD-9-CM and ICD-10) and were augmented through lexical search and semantic exploration of the OHDSI standardized vocabularies.
Cohort definitions and associated conceptsets were then reviewed by a clinical adjudicator.
Cohort definitions were developed using the OHDSI open-source platform ATLAS (https://github.com/OHDSI/atlas).
These definitions were initially executed across 7 databases (CCAE, MDCR, MDCD, Optum, Panther, JMDC, IMS Germany) to identify qualifying patients.
Because the databases used in this study do not all consistently contain laboratory values, outcomes involving electrolyte imbalance (hypokalemia, hypomagnesemia, hyponatremia) were defined by diagnosis records only.
These cohorts were used to produce characterization of outcome incidence, stratified by age decile, gender, and index year, to determine if outcomes were being observed in a frequency that was generally consistent across data sources and sufficiently aligned with epidemiologic expectations of the disease.
No source record verification or other methods for validation was performed.       

```{r, echo=FALSE}
tab <- read.csv("CohortDefinitionTable.csv")
tab$name <- paste0("[", tab$name, "](", tab$url, ")")
tab <- tab %>% select(name, description, citations)
names(tab) <- c("Phenotype", "Logical description", "Supporting references")
knitr::kable(tab)
```

## References
