---
csl: pnas.csl

output:
  pdf_document:
    fig_caption: yes
  html_document: default
bibliography: bibliography.bib
params:
  databaseId: "MDCD"
  targetId: 739138
  comparatorId: 715259
  outcomeId: 18
  setTitle: "A Comparison of Sertraline to Duloxetine for the Risk of Stroke in the MDCD Database."
title: "`r params$setTitle`"
---


```{r, echo=FALSE, message=FALSE, comment=FALSE, warning=FALSE, results='hide'}
library(DatabaseConnector)
library(knitr)
library(kableExtra)
source("DataPulls.R")
source("PlotsAndTables.R")
options(knitr.kable.NA = '')

# params <- list(databaseId = "MDCR",
#                targetId = 739138,
#                comparatorId =  715259,
#                outcomeId = 18)

useStoredObject <- FALSE

if (!useStoredObject) {
    connectionDetails <- createConnectionDetails(dbms = "postgresql",
                                                 server = "localhost/ohdsi",
                                                 user = "postgres",
                                                 password = Sys.getenv("pwPostgres"),
                                                 schema = "legend")
    connection <- connect(connectionDetails)
    targetName <- getExposureName(connection, params$targetId)
    comparatorName <- getExposureName(connection, params$comparatorId)
    outcomeName <- getOutcomeName(connection, params$outcomeId)
    analyses <- getAnalyses(connection)
    mainResults <- getMainResults(connection,
                                  targetIds = params$targetId,
                                  comparatorIds = params$comparatorId,
                                  outcomeIds = params$outcomeId,
                                  databaseIds = params$databaseId,
                                  analysisIds = c(1, 2, 3, 4))
    
    controlResults <- getControlResults(connection,
                                        targetId = params$targetId,
                                        comparatorId = params$comparatorId,
                                        analysisId = 1,
                                        databaseId = params$databaseId)

    attrition <- getAttrition(connection,
                              targetId = params$targetId,
                              comparatorId = params$comparatorId,
                              outcomeId = params$outcomeId,
                              analysisId = 1,
                              databaseId = params$databaseId)

    balance <- getCovariateBalance(connection,
                                   targetId = params$targetId,
                                   comparatorId = params$comparatorId,
                                   databaseId = params$databaseId)

    ps <- getPs(connection,
                targetId = params$targetId,
                comparatorId = params$comparatorId,
                databaseId = params$databaseId)

    kaplanMeier <- getKaplanMeier(connection,
                                  targetId = params$targetId,
                                  comparatorId = params$comparatorId,
                                  outcomeId = params$outcomeId,
                                  databaseId = params$databaseId,
                                  analysisId = 2)
} else {
    load("paperData.rda")
}
```

\centerline{Martijn J. Schuemie$^{1,2}$}
\centerline{Marc A. Suchard$^{1,3,4,5}$}
\centerline{George M. Hripcsak$^{1,6}$}
\centerline{Patrick B. Ryan$^{1,2,6}$}
\centerline{David Madigan$^{1,7}$}

$^{1}$ Observational Health Data Sciences and Informatics, New York, NY
$^{2}$ Janssen Research & Development, Titusville, NJ
$^{3}$ Department of Biomathematics, University of Califoria, Los Angeles, CA
$^{4}$ Department of Biostatistics, University of Califoria, Los Angeles, CA
$^{5}$ Department of Human Genetics, University of Califoria, Los Angeles, CA
$^{6}$ Department of Biomedical Informatics, Columbia University, New York, NY
$^{7}$ Department of Statistics, Columbia University, New York, NY

Corresponding author: Martijn J. Schuemie, Janssen R&D, 1125 Trenton Harbourton Road, Titusville, NJ, 08560, Phone: +31 631793897, schuemie@ohdsi.org

# Abstract

To do

# Introduction

This is a very important study. Here's a really cool paper @pmid23900808.

# Methods

# Results

```{r, echo = FALSE, fig.width=6, fig.height=7.5, out.width = '50%', fig.align='center'}
drawAttritionDiagram(attrition, targetName, comparatorName)
```
**Figure 1**. Attrition diagram.


**Table 1**. Select population characteristics
```{r, echo = FALSE}
table <- prepareTable1(balance)
# Remove 1st header, with add back later with merged columns:
header <- as.character(table[1, ])
header[header == "1"] <- ""
table <- table[-1, ]
# Indentation needs to be made explicit (not by leading spaces):
needIndent <- which(substr(table[, 1], 1, 1) == " ")

kable_styling(add_indent(add_header_above(kable(table, "latex", 
                                                booktabs = TRUE, 
                                                longtable = TRUE,
                                                row.names = FALSE, 
                                                col.names = header,
                                                linesep = "",
                                                align = c("l", "r", "r", "r", "r", "r", "r")), 
                                          c("", "Before stratification" = 3, "After stratification" = 3)),
                         needIndent),
              font_size = 7,
              latex_options = c("HOLD_position", "repeat_header"))
           
```

```{r, echo = FALSE, fig.width=5, fig.height=3.5, out.width = '50%', fig.align='center'}
plotPs(ps, targetName, comparatorName)
```
**Figure 2**. Preference score distribution. The preference score is a transformation of the propensity score that adjusts for differences in the sizes of the two treatment groups. A higher overlap indicates subjects in the two groups were more similar in terms of their predicted probability of receiving one treatment over the other.

```{r, echo = FALSE, fig.width=4, fig.height=4, out.width = '50%', fig.align='center', warning=FALSE}
CohortMethod::plotCovariateBalanceScatterPlot(balance, beforeLabel = "Before stratification", afterLabel = "After stratification")
```
**Figure 3**. Covariate balance before and after stratification. Each dot represents the standardizes difference of means for a single covariate before and after stratification on the propensity score. 

```{r, echo = FALSE, fig.width=15, fig.height=2.5, out.width = '100%', fig.align='center', warning=FALSE}
EmpiricalCalibration::plotCiCalibrationEffect(logRr = controlResults$logRr,
                                              seLogRr = controlResults$seLogRr,
                                              trueLogRr = log(controlResults$effectSize),
                                              xLabel = "Hazard ratio")
```
**Figure 4**. Systematic error 


**Table 2**. Hazard ratios, 95% confidence intervals, uncalibrated and empirically calibrated, for various analyses.
```{r, echo = FALSE}
table <- mainResults
table$hr <- sprintf("%.2f (%.2f - %.2f)", mainResults$rr, mainResults$ci95lb, mainResults$ci95ub)
table$p <- sprintf("%.2f", table$p)
table$calHr <- sprintf("%.2f (%.2f - %.2f)", mainResults$calibratedRr, mainResults$calibratedCi95Lb, mainResults$calibratedCi95Ub)
table$calibratedP <- sprintf("%.2f", table$calibratedP)
table <- merge(table, analyses)
table <- table[, c("description", "hr", "p", "calHr", "calibratedP")]
colnames(table) <- c("Analysis", "HR (95% CI)", "P", "Cal. HR (95% CI)", "Cal. p")
kable_styling(kable(table, "latex", 
                    booktabs = TRUE, 
                    longtable = FALSE,
                    row.names = FALSE, 
                    linesep = ""),
              font_size = 8,
              latex_options = c("HOLD_position"))
```

```{r, echo = FALSE, fig.width=7, fig.height=5, out.width = '100%', fig.align='center', results='hide'}
plotKaplanMeier(kaplanMeier, targetName, comparatorName)
```
**Figure 3**. Kaplan Meier plot, showing survival as a function of time. This plot
      is adjusted for the propensity score stratification: The target curve (<em>`r targetName`</em>) shows the actual observed survival. The
      comparator curve (<em>`r comparatorName`</em>) applies reweighting to approximate the counterfactual of what the target survival
      would look like had the target cohort been exposed to the comparator instead. The shaded area denotes
      the 95 percent confidence interval.


# Conclusions

# References

```{r, echo=FALSE, message=FALSE, comment=FALSE, warning=FALSE, results='hide'}
if (!useStoredObject) {
  DatabaseConnector::disconnect(connection)
}
```
